---
title: "Instacart Buyer Analysis"
author: "Sandeep Anand"
date: "August 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Get the data

```{r Data2, eval=TRUE, warning=FALSE, echo=FALSE}
path1 <- "C:/Public/Instacart/"
names <- character(0)
for (j in list.files(path = path1, pattern = "\\.csv$")) 
  {
    names<-c(names,unlist(strsplit(j,"\\."))[1])
}
names
```

## Get it into a data.table

```{r Data3, eval=TRUE, warning=FALSE, echo=FALSE}
library(data.table)
filenames<-list.files(path = path1, pattern = "*.csv$", full.names = TRUE)
ldf <- lapply(filenames, fread)
filenames
```

## Assignment here

```{r assignment, eval=TRUE, warning=FALSE, echo=FALSE}
for(i in 1:(length(names)-1)){
  nam <- paste(names[i], "dt", sep = ".")
  assign(nam, ldf[[i]])
}
```

## Assignments to be used for Data-frame names:
  - "aisles.dt"
  - "departments.dt"
  - "order_products__prior.dt"
  - "order_products__train.dt"
  - "orders.dt" 

```{r Orders.df Proc, eval=TRUE, warning=FALSE, echo=FALSE}
order_train.dt<-orders.dt[orders.dt$eval_set=="train",]
order_prior.dt<-orders.dt[orders.dt$eval_set=="prior",]
order_test.dt<-orders.dt[orders.dt$eval_set=="test",]

head(orders.dt)

head(order_products__prior.dt)

head(order_prior.dt)
```

## Applying xgboost and post processing the data using data frames only

 - The same can be done using data table also , but chose df instead
 - xgboost used all numerical vectors only so, it is important that all features to be used are numerical
 - understand the usage of inner_join , semi_join and left_join 
 - Formation of a new df using orders.dt and prior orderered products

```{r post processing to apply xgboost, message=FALSE, warning=FALSE, echo=FALSE, eval=TRUE}
library(data.table)
library(dplyr)
library(tidyr)

head(order_train.dt)
head(order_test.dt)

# Reshaping the data

aisles.dt$aisle <- as.factor(aisles.dt$aisle)
departments.dt$department <- as.factor(departments.dt$department)
orders.dt$eval_set <- as.factor(orders.dt$eval_set)
products.dt$product_name <- as.factor(products.dt$product_name)

products.dt <- products.dt %>% inner_join(aisles.dt) %>% inner_join(departments.dt) %>% select(-aisle_id, -department_id)

rm(aisles.dt, departments.dt)

head(products.dt)

# All the Order Ids are unique in Oders.csv so, the below match can be done 
order_products__train.dt$user_id <- orders.dt$user_id[match(order_products__train.dt$order_id, orders.dt$order_id)]

order_products <- orders.dt %>% inner_join(order_products__prior.dt, by="order_id")

#just cheking what we are getting here
head(order_products__train.dt,20)
head(order_products,200)
#order_prior.dt[which(order_products$order_id==2539329)]

rm(order_products__prior.dt)
gc()
```

## Products
  - creating Products feature set

```{r Products, message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}

# multiheirarchical orders
prd <- order_products %>% arrange(user_id, order_number, product_id) %>% group_by(user_id, product_id) %>% 
        mutate(product_time = row_number()) %>% ungroup() %>%
        group_by(product_id) %>%
        summarise(
          prod_orders = n(),
          prod_reorders = sum(reordered),
          prod_first_orders = sum(product_time == 1),
          prod_second_orders = sum(product_time == 2)
        )
# Everything is Prior here as we got this using inner join with order_products__prior.dt
head(prd)


prd$prod_reorder_probability <- prd$prod_second_orders/prd$prod_first_orders
prd$prod_reorder_times <- 1+ prd$prod_reorders / prd$prod_first_orders
prd$prod_reorder_ratio  <- prd$prod_reorders / prd$prod_orders

prd <- prd %>% select(-prod_reorders, -prod_first_orders, -prod_second_orders)

# We Probably have everything here with Products, so will not require this dt anymore
rm(products.dt)
gc()
```

## users 
  - Getting the user feture set here
```{r Users, message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE}

# Everything we are dealing with is prior only , all the data collected so for user id also we will need the same 
head(orders.dt,10)
users <- orders.dt %>% 
          filter(eval_set == "prior") %>%
          group_by(user_id) %>%
          summarise(
            user_orders = max(order_number),
            user_period = sum(days_since_prior_order),
            user_mean_days_since_prior = mean(days_since_prior_order, na.rm=T)
          )
      


us <- order_products %>% 
      group_by(user_id) %>%
      summarise(
        user_total_products = n(),
        user_reorder_ratio = sum(reordered == 1)/sum(order_number>1),
        user_distinct_products = n_distinct(product_id)
      )

head(users,10)
head(us, 10)

users <- users %>% inner_join(us)
users$user_average_basket <- users$user_total_products/users$user_orders


us <- orders.dt %>% 
    filter(eval_set != "prior") %>%
    select(user_id, order_id, eval_set, time_since_last_order = days_since_prior_order)

users <- users %>% inner_join(us)

head(users,10)

rm(us)
gc()
```


## database 
  - Creating the entire dataset with feature sets that can be directly used with xgboost

```{r DB, message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE}

data<-NULL
data <- order_products %>% 
        group_by(user_id, product_id) %>%
        summarise(
          up_orders = n(),
          up_first_order = min(order_number),
          up_last_order = max(order_number),
          up_average_cart_position = mean(add_to_cart_order)
        )

head(data,50)

rm(order_products, orders)


# Remember - Each order has multiple products here so "user_orders" is the total orders for each user

data <- data %>% 
       inner_join(prd, by = "product_id") %>%
       inner_join(users, by = "user_id")

data$up_order_rate <- data$up_orders/data$user_orders
data$up_orders_since_last_order<-data$user_orders - data$up_last_order # orders the user ordered after this user , product combination
data$up_order_rate_since_first_order <- data$up_orders / (data$user_orders - data$up_first_order + 1) # up_orders is basically the total number of times the particular product appeas for that user for all the orders for that user

data <- data %>% left_join(order_products__train.dt %>% select(user_id, product_id, reordered), by = c("user_id", "product_id"))

rm(order_products__train.dt, prd, users)
gc()

```

## Train/Test Datasets

```{r Train/test ,message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE}

head(data,100)
train <- as.data.frame(data[data$eval_set == "train",])

train$eval_set <- NULL
train$user_id <- NULL
train$product_id <- NULL
train$order_id <- NULL
train$reordered[is.na(train$reordered)] <- 0

test <- as.data.frame(data[data$eval_set == "test",])

test$eval_set <- NULL
test$user_id <- NULL
test$reordered <- NULL

rm(data)
gc()

```

## Light GBM - or any other xgboost technique
 - Using 1st set of parameters gives me a F1 score = 0.363

```{r lightgbm ,message=FALSE, warning=FALSE, echo=TRUE, eval=FALSE}
library(xgboost)

# 1st try = 0.4, 2nd try = 0.1
subtrain <- train %>% sample_frac(0.10) # Sample size is up for discussion you can choose as you find optimal


# This is the model
xgb.train <- xgboost(data = as.matrix(subtrain %>% select(-reordered)), 
                    label = subtrain$reordered, 
                    max_depth = 2, 
                    eta = 0.1, 
                    nthread = 2, 
                    nrounds = 100,
                    max_depth = 20,
                    subsample = 0.5,
                    colsample_bytree = 0.5,
                    eval_metric = "logloss",
                    objective = "binary:logistic")
# Get Feature importance
xgb.feature.imp <- xgb.importance(model = xgb.train, colnames(subtrain))

# Get test predictions
testdata <- as.matrix(test %>% select(-order_id, -product_id))

test$reordered <- predict(xgb.train, testdata)

test$reordered <- (test$reordered > 0.21) * 1

head(test,100)

submission <- test %>% 
              filter(reordered == 1) %>% 
              group_by(order_id) %>%
              summarise(
                products = paste(product_id, collapse = " ")
              )

missing <- data.frame(
          order_id = unique(test$order_id[!test$order_id %in% submission$order_id]),
          products = "None"
)

submission <- submission %>% bind_rows(missing) %>% arrange(order_id)

write.csv(submission, file = "submission5.csv", row.names = F)
```

## more attempts
  - Attempts to modify the current parameters to make a better submission

```{r xgboost-2, warning=FALSE, message=FALSE, echo=TRUE, eval=TRUE}

# Model -------------------------------------------------------------------
library(xgboost)

#params <- list(
#  "objective"           = "reg:logistic",
#  "eval_metric"         = "logloss",
#  "eta"                 = 0.1,
#  "max_depth"           = 6,
#  "min_child_weight"    = 10,
#  "gamma"               = 0.70,
#  "subsample"           = 0.76,
#  "colsample_bytree"    = 0.95,
#  "alpha"               = 2e-05,
#  "lambda"              = 10
#)

params <- list (
  "objective"           = "binary:logistic",
  "eval_metric"         = "logloss",
  "eta"                 = 0.1,
  "max_depth"           = 6,
  "min_child_weight"    = 25,
  "gamma"               = 0.70,
  "subsample"           = 0.76,
  "colsample_bytree"    = 0.95,
  "alpha"               = 2e-05,
  "lambda"              = 25  

)

# 1st try = 0.1, 2nd try = 0.33 , 4th try = 0.40
subtrain <- train %>% sample_frac(0.40) # Sample size is up for discussion you can choose as you find optimal

X<-xgb.DMatrix(as.matrix(subtrain %>% select(-reordered)), label=subtrain$reordered)
model<-xgboost(data = X, params = params, nrounds = 100)

importance <- xgb.importance(colnames(X), model = model)
xgb.ggplot.importance(importance)

rm(X, importance, subtrain)
gc()

#Apply Model
X<-xgb.DMatrix(as.matrix(test %>% select(-order_id, -product_id)))
test$reordered <- predict(model, X)

test$reordered <- (test$reordered > 0.21)*1

submission<-NULL

submission <- test %>% 
              filter(reordered == 1) %>%
              group_by(order_id) %>%
              summarize(
                products = paste(product_id, collapse = " ")
              )

# getting the product Ids 

ps1<-"C:/Public/Instacart/submission4.csv"
solold.dt <- fread(ps1)

missing<-NULL

missing <- data.frame(
  order_id = unique(test$order_id[!test$order_id %in% submission$order_id]),
  products = "None"
)

# The below older solution is not useful
#temp<-missing %>% inner_join(solold.dt, by = "order_id")
#missing$products <- temp$products

submission <- submission %>% bind_rows(missing) %>% arrange(order_id)

write.csv(submission, file = "submission7.csv", row.names = F)
```


